---
title: "p8105_hw2_cc5604"
author: "Changshen Chen"
date: 2026-10-04
output: github_document
---

# Problem 1
## Loading necessary package
```{r}
library(tidyverse)
```

## Reading and cleaning the designated file
```{r}
sub_df = 
  read.csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |> 
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) |> 
  mutate(entry = case_match(entry, "YES" ~ TRUE, "NO" ~ FALSE))
```

## Dataset description
After cleaning, the dataset contains line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance information. For the cleaning steps, I first used `clean_names()` function to convert the raw column names into standardized strings. Then, I used the `select()` function to keep only the variables of interest listed above. Lastly, I used the `mutate()` function to convert the entry variable from character (YES vs NO) to a logical variable. The resulting dataset now has `r nrow(sub_df)` rows and `r ncol(sub_df)` columns. These data are not tidy because the route number served is stored in the column names instead of in a single variable.

## Answering the questions
1) There are `r nrow(distinct(sub_df, line, station_name))` distinct stations.

2) `r nrow(sub_df |> filter(ada == "TRUE") |> distinct(line, station_name))` stations are ADA compliant.

3) `r sub_df |> filter(vending == "NO") |> pull(entry) |> mean()*100`% of station entrances / exits without vending allow entrance.

## Reformatting the dataset
```{r}
tidy_df = 
  sub_df |> 
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  )
```

## Answering the questions
1) `r nrow(tidy_df |> filter(route_name == "A") |> distinct(line, station_name))` distinct stations serve the A train.
2) Of the stations that serve the A train, `r nrow(tidy_df |> filter(route_name == "A") |>  filter(ada == "TRUE") |> distinct(line, station_name))` are ADA compliant.

# Problem 2
## Reading and cleaning the designated file
```{r}
mr_df = 
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N709") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster)) |>
  mutate(sports_balls = as.integer(round(sports_balls)))

prof_df = 
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M134") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster))

gwy_df =
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel", range = "A2:L351") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster))
```

## Combining the datasets
```{r}
mr_df = mutate(mr_df, trash_wheel = "Mr. Trash Wheel", year = as.numeric(year))
prof_df = mutate(prof_df, trash_wheel = "Professor Trash Wheel")
gwy_df = mutate(gwy_df, trash_wheel = "Gwynns Falls Trash Wheel")

trash_df = bind_rows(mr_df, prof_df, gwy_df)

trash_df
```

## Dataset description
The combined dataset contains information on dumpsters from Mr. Trash Wheel, Professor Trash Wheel, and Gwynns Falls Trash Wheel with rows that do not include dumpster-specific data omitted. The resulting dataset has `r nrow(trash_df)` rows and `r ncol(trash_df)` columns, which meaning there are `r nrow(trash_df)` observations and `r ncol(trash_df)` variables. I added an extra variable named `trash_wheel` to keep track of which Trash Wheel is which when binding the three datasets together. The key variables include time information such as `month` and `year`, trash amount information such as `weight_tons` and `volume_cubic_yards`, and specific trash record including `cigarette_butts`, `plastic_bottles`, and `plastic_bags`. Some variables are not included in all wheels, for example `sports_balls` is only available for Mr. Trash Wheel.

## Answering the questions
```{r}
weight =
  trash_df|> 
  filter(trash_wheel == "Professor Trash Wheel") |> 
  pull(weight_tons) |> 
  sum()

cig = 
  trash_df |> 
  filter(trash_wheel == "Gwynns Falls Trash Wheel") |> 
  filter(month == "June") |> 
  filter(year == "2022") |> 
  pull(cigarette_butts) |> 
  sum()
```

For the available data, `r weight` tons of trash were collected by Professor Trash Wheel. `r cig` cigarette butts were collected by Gwynnda in June of 2022.

# Problem 3
## Reading and cleaning the baseline file
```{r}
baseline_df = 
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1, na = c(".")) |> 
  janitor::clean_names()|>
  mutate(
    sex = case_match(
      sex, 
      1 ~ "male", 
      0 ~ "female"
      ),
    apoe4 = case_match(
      apoe4, 
      1 ~ "carrier", 
      0 ~ "non-carrier"
      )
  ) 

clean_baseline_df = 
  baseline_df |>
  filter(is.na(age_at_onset) | age_at_onset > current_age)

n_women = nrow(clean_baseline_df |> filter(sex == "female"))
n_women_carrier = nrow(clean_baseline_df |> filter(sex == "female", apoe4 == "carrier") )
women_prop = (n_women_carrier / n_women) *100
```

The dataset was imported using the `read_csv()` function. I used `skip = 1` arugument to omit the non-data entry and capture the real header. The `na = c(".")` argument is used because a missing age at onset is written as a period. For the cleaning steps, I first used `clean_names()` function to convert the raw column names into standardized strings. Then, I used the `mutate()` function to convert the sex and apoe4 variable from numeric to character labels adapted from the data description row. Lastly, I used the `filter()` function to remove participants who do not have MCI at baseline. The dataset contains participant id, current age, sex, education, apoe4 carrier, and age at onset informtion. `r nrow(baseline_df)` participants were recruited, `r nrow(clean_baseline_df)` participants met the stated inclusion criteria, and `r nrow(clean_baseline_df |> filter(!is.na(age_at_onset)))` developed MCI. The average baseline age is `r mean(clean_baseline_df$current_age)`. `r women_prop`% of women in the study are APOE4 carriers.

## Reading and cleaning the amyloid file
```{r}
amyloid_df = 
  read_csv("data/data_mci/mci_amyloid.csv", skip = 1, na = c("NA", "Na")) |> 
  janitor::clean_names() |>
  pivot_longer(
    baseline:time_8,
    names_to = "time",
    values_to = "biomarker_values"
  )
```

For the amyloid data, I used `skip = 1` argument to omit the non-data entry and capture the real header. The `na = c("NA", "Na")` argument is used because the na values are represented by NA or Na in the raw dataset. For the cleaning steps, I first used `clean_names()` function to convert the raw column names into standardized strings. Then, I used `pivot_longer()` to store the five time-related features in a single variable with another column matching the biomarker values to the corresponding time of visit. The cleaned dataset now has `r nrow(amyloid_df)` rows and `r ncol(amyloid_df)` columns. It contains study id, time, and biomarker value information.

