---
title: "p8105_hw2_cc5604"
author: "Changshen Chen"
date: 2026-10-04
output: github_document
---

# Problem 1
## Loading necessary package
```{r}
library(tidyverse)
```
## Reading and cleaning the designated file
```{r}
sub_df = 
  read.csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |> 
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) |> 
  mutate(entry = case_match(entry, "YES" ~ TRUE, "NO" ~ FALSE))
```

## Dataset description
After cleaning, the dataset contains line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance information. For the cleaning steps, I first used `clean_names()` function to convert the raw column names into standardized strings. Then, I used the `select()` function to keep only the variables of interest listed above. Lastly, I used the `mutate()` function to convert the entry variable from character (YES vs NO) to a logical variable. The resulting dataset now has `r nrow(sub_df)` rows and `r ncol(sub_df)` columns. These data are not tidy because the route number served is stored in the column names instead of in a single variable.

## Answering the questions
1) There are `r nrow(distinct(sub_df, line, station_name))` distinct stations.

2) `r nrow(sub_df |> filter(ada == "TRUE") |> distinct(line, station_name))` stations are ADA compliant.

3) `r sub_df |> filter(vending == "NO") |> pull(entry) |> mean()*100`% of station entrances / exits without vending allow entrance.

## Reformatting the dataset
```{r}
tidy_df = 
  sub_df |> 
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  )
```

## Answering the questions
1) `r nrow(tidy_df |> filter(route_name == "A") |> distinct(line, station_name))` distinct stations serve the A train.
2) Of the stations that serve the A train, `r nrow(tidy_df |> filter(route_name == "A") |>  filter(ada == "TRUE") |> distinct(line, station_name))` are ADA compliant.

# Problem 2
## Reading and cleaning the designated file
```{r}
mr_df = 
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N709") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster)) |>
  mutate(sports_balls = as.integer(round(sports_balls)))

prof_df = 
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M134") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster))

gwy_df =
  readxl::read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel", range = "A2:L351") |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster))
```





